Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages

### 2021-03-06 manuscript/bef_stability/manuscript.Rmd
Reviewed according to Elisa's comments

---
title: Species richness and food-web structure jointly drive community biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault, Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

Biodiversity-ecosystem functioning (BEF) and food-web complexity-stability
relationships are central to ecology. However, they remain largely untested in
natural context. Here we study the links among environmental conditions,
richness, food-web structure, annual biomass and its temporal stability using a
standardized monitoring dataset of 99 stream fish communities spanning from 1995
to 2018. We first reveal that both richness and average trophic level are
positively related to annual biomass, with effects of similar strength. Second,
we show that food-web structure affects community stability, with a positive
effect of mean trophic level, and contrary to expectation, a destabilizing
effect of species richness. Finally, we find that environmental conditions
affect both biomass and its stability via effects on richness and network
structure. Strikingly, effects on stability are mediated by changes in
population stability rather than synchrony, which contrasts with results from
single-trophic communities and might be characteristic of multi-trophic
communities.

@@

Biodiversity affects biomass production [@tilman_productivity_1996;
@cardinale_species_2002; @grace_integrative_2016], distribution
[@thebault_relationship_2006] and stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006]. Indeed species richness
generally increases primary productivity and total biomass in experimental plant
communities [@tilman_productivity_1996; @grace_integrative_2016;
@hector_plant_1999; @tilman_biodiversity_2014; @grace_integrative_2016], stream
feeders, [@cardinale_species_2002] and aquatic bacterial consumers
[@pennekamp_biodiversity_2018]. In parallel, such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@houlahan_negative_2017; @olivier_urbanization_2020].
Despite the diversity of ecosystems and taxa studied, the relationship between
diversity and ecosystem functioning remains mostly focused on single trophic
levels, ignoring the effects of food-web structure. This gap is even more
striking for the relationship between diversity and temporal stability , in
spite of long-standing interest for the links between food-web complexity and
stability in theoretical ecology [e.g.  @may_will_1972; @dunne_network_2006;
@neutel_reconciling_2007; @duffy_why_2009; @stouffer_compartmentalization_2011].

Overall, the evidence from theoretical models suggests that both species
richness and food-web structure should modulate ecosystem properties such as
community biomass and its stability. Complementarity in the use of
resources is important to understand the effects of species richness in food-webs
[@thebault_food-web_2003; @poisot_trophic_2013; @schneider_animal_2016], but it
also depends on food-webs connectance, i.e. its level of generalism
[@thebault_food-web_2003; @poisot_trophic_2013].

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020],
asynchrony in population temporal variations also explains a significant part
of the stabilizing effects of species richness on community biomass in food-webs
[@thebault_trophic_2005]. Indeed, a low synchrony in population fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a population is compensated by the increase of another. However,
higher food-web connectance can counterbalance the positive effects of species
richness on temporal stability of community biomass, because it may increase
competition between consumers and then leads to a positive relationship between
species richness and population temporal variability
[@thebault_trophic_2005]. In addition, the presence of species at higher trophic
levels increases total primary production, through stronger top-down effects
associated with higher nutrient uptake [@wang_biodiversity_2018]. The addition of predators has
also cascading effects on both population variability and
synchrony at lower trophic levels [@teng_dynamics_2004;
@shanafelt_stability_2018]. Overall, these theoretical studies show that the
effects of species richness and food-web structure, such as connectance and
average trophic level, are potentially conflicting or synergistic in such a way
that they need to be addressed jointly.

The lack of studies that consider jointly effects of diversity and food-web
structure on ecosystem functions can be explained by the challenges associated
with manipulating both species richness and food-web complexity in an
experimental context [e.g. @sander_ecological_2017] and by the fact that
describing trophic interactions between species is highly labour intensive.
These challenges are even more acute for the study of the stability of ecosystem
properties, which additionally requires longer-term perspectives. In this
context, recent developments to predict species trophic interactions
[@beckerman_foraging_2006; @gravel_inferring_2013; @portalier_foraging_2019],
combined with observational long-term surveys of ecological communities, can
offer the opportunity to investigate the links between diversity, food-web
structure and the stability of total community biomass in natural settings.

Such approach based on data from natural communities can, in addition, give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
population abundances, community biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus community biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
food-web structure, as larger ecosystems host food-webs with higher
diversity and trophic levels [@mchugh_dual_2010] whereas temperature can affect
food-web diversity and structure in contrasted ways depending on the ecosystem
studied [@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. In natural
communities, the relationships among species richness, community biomass and its
temporal stability might also arise from effects of the environment on both
diversity and ecosystem functions [@loreau_biodiversity_2001; @he_global_2013;
@grace_integrative_2016].

@@

Because their trophic interactions are well known, stream fish communities
constitute a good model to study the effects of network structure on total
biomass and its temporal stability. A substantial body of literature
supports that trophic interactions can be inferred through body size and diet
informations [@jennings_weak_2001; @beckerman_foraging_2006;
@gravel_inferring_2013; @brose_predator_2019; @portalier_foraging_2019].

@@

```{r, cache = FALSE}
pca_caption <- paste0(
  "(B) Rotated axis of PCA on environmental variables. The five components
  explained,", tot_var_expl_pca, "% of the variance (",
  paste0(names(prop_var_expl_pca), ":", prop_var_expl_pca, "%"), ").",
  " Axis 1 was positively related to the stream size, axis 2 was related to
  the stream slope and altitude and negatively related to average water temperature and
  BOD."
  )

@@

We selected the stations that had been sampled ten years or more. As the concept
of community variability referred to variability around an equilibrium, which is
rarely observed or difficult to characterize in natural communities, only
stations with no temporal trends in community biomass were selected (detailed in
appendix B, part 2). However, the results were robust to the relaxation of this
hypothesis (Fig. S7). These selection steps resulted in 99 stations distributed
over seven hydrographic basins (Fig. \@ref(fig:matmet)A).

@@

The body size of all fish individuals was not recorded when they were too
numerous (Fig. S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution (Fig. S1, appendix B, part 1.2). The body mass of each
individual was then estimated following the results of two meta-analyses on
fishes [@lleonart_removing_2000; @froese_cube_2006], using an allometric law
$B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams, $l$ the body
size in millimeters and $i$ the individual. Population biomass and community biomass at
each sampling event were then calculated by summing individual body masses. To
account for the variation in sampling effort among sites, population biomass and
community biomass were corrected by the surface sampled (range `r range_surface[1]` -
`r range_surface[2]` square metres), and expressed in grams per square meter
($gm^{-2}$). In the analysis, the community biomass of a site was defined as
the median of the community biomass across sampling years.

@@

Temporal biomass stability was defined as the inverse of biomass variability.
Variability was computed as Coefficient of Variation (CV) of the biomass over
time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and $\mu_i$ being
respectively the standard deviation and the temporal average of the annual total
biomass of the station $i$). Then, temporal stability ($S_i$) was equal to $CV^{-1} =
\mu \sigma^{-1}$.

We described the mechanisms driving the temporal stability of biomass,
decomposing the $CV$ of the community biomass ($CV_{com}$) into a
synchrony ($\phi$) component and a weighted average species $CV$ ($\overline{CV_{sp}}$)
component according to @thibaut_understanding_2013:

@@

with $CV_{com, i}$ being the CV of the community biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average CV at the species-level in the station $i$, weighted by the relative biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
community biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_{k,i}}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

@@

In further analysis, we linearized the mathematical relationship linking
temporal stability to $\overline{CV_{sp, i}}$ and $\phi$ by taking the log of
temporal stability:

\begin{equation} 
  \begin{aligned}
    S_i &= (\overline{CV_{sp, i}} * \phi_i^\frac{1}{2})^{-1} \\
    \log S_i &= - \log (\overline{CV_{sp, i}} * \phi_i^\frac{1}{2}) \\
    \log S_i &= - \log (\overline{CV_{sp, i}} - \log \phi_i^\frac{1}{2} \\
    \log S_i &= - \log \overline{CV_{sp, i}} - \frac{1}{2} \log \phi_i \\
  \end{aligned}
  (\#eq:stab)
\end{equation}

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all fishes and the resources (Fig.
\@ref(fig:matmet)B), then (2) determine network of a given community by
extracting the metaweb according to the fishes present at a given
site and sampling event (Fig. \@ref(fig:matmet)C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, larvae and small juveniles feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (appendix B, Fig. S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Fig. \@ref(fig:matmet)B, C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

@@

For each trophic network, corresponding to a particular site and year, we
computed the connectance and mean trophic level. Connectance describes the
level of generalism in the network and is calculated as the actual number of
interactions divided by the number of interactions if all species interact
($C = L / N^2$, L and N being resp. the number of links and of nodes).
The trophic level of the nodes was computed as one added to the average trophic
level of its food item. The mean trophic level $\overline{T}$ was 
weighted by the biomass of each trophic species (excluding resource nodes for
which we had no biomass information), so $\overline{T} = \sum_i^N{B_iT_i \times
(\sum_i^N{B_i})^{-1}}$, with $T_i$ and $B_i$ being respectively the trophic
level and the biomass of the node $i$. For each site, the median of connectance
and average trophic level over time was used in subsequent analysis.

@@

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and temporal CV values of water temperature, flow, BOD, river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (appendix B, Fig. S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (appendix B, Fig. S4). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, average BOD and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

@@

We analyzed the relationships among environment, species richness, food-web
structure, community biomass and its temporal stability with structural equation
models as they enable the assessment of potential indirect and direct causal
relationships among multiple variables [@grace_structural_2008;
@grace_causal_2014; @lefcheck_piecewisesem_2016; @grace_integrative_2016].  We
defined a structural equation model explaining the median annual community
biomass as a function of the two synthetic environmental variables, species
richness and food-web structure [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and food-web
structure directly affect community biomass (Fig. \@ref(fig:sem-plot)A) and that
species richness affected food-web structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

We used the same model structure to assess the drivers of temporal stability
of community biomass (Fig.  \@ref(fig:sem-plot)C). Instead of the
annual community biomass, we modelled the synchrony and $\overline{CV_{sp}}$ which
in turn determined the temporal stability of community biomass.

@@

The structural equation models explained a large part of the variation
of species richness, connectance, synchrony, $\overline{CV_{sp}}$ and total
biomass (resp. $R^2 = .43$, $.38$, $.46$, $.47$ and $.32$) but less so for mean
trophic level ($R^2 = .11$, Fig. \@ref(fig:sem-plot)B and D). We found a
negative relationship between species richness and mean trophic level ($r_\delta
= `r rich_on_tlvl`$, $r_\delta$ meaning standardized slope) and connectance
($r_\delta = `r rich_on_ct`$, Fig.  \@ref(fig:sem-plot)B and D), meaning that a
part of the effects of species richness on community biomass and its temporal
stability went through food-web structure.

@@

Species richness and mean trophic level had a significant positive direct effect
on community biomass ($r_\delta = `r rich_on_bm$direct`$ and 
$r_\delta = `r tlvl_on_bm`$ resp., Fig. \@ref(fig:sem-plot)B, Table
\@ref(tab:indir-sem)), while connectance had no direct effect.

The two PCA axes synthesising environmental variables had no significant direct
effects on community biomass, but indirect (Table \@ref(tab:indir-sem)). The first
axis, related to the stream size, was positively linked to community biomass
($r_\delta = `r env_on_bm$total["log_RC1"]`$, Table \@ref(tab:indir-sem)),
through both species richness and mean trophic level ($r_\delta = 
`r env_on_bm$via_richness["log_RC1"]`$ and $`r env_on_bm$via_tlvl["log_RC1"]`$
resp., Table \@ref(tab:indir-sem)). The second axis, related to the average
annual temperature was positively related
to community biomass ($r_\delta = `r env_on_bm$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)), mainly through species richness ($r_\delta = 
`r env_on_bm$via_richness["log_RC2"]`$, Table \@ref(tab:indir-sem)). The total effect
of PCA1, PCA2 and species richness on community biomass were of the same magnitude.

@@

By its effects on synchrony and $\overline{CV_{sp}}$, which respectively
measures the level of compensatory dynamics among population and the average
population variability, species richness had a
negative total effect on biomass stability ($r_\delta = `r rich_on_stab$total`$,
Table \@ref(tab:indir-sem)). This effect resulted from two opposite effects. On
the one hand, species richness had a direct negative effect on synchrony
($r_\delta = `r rich_on_sync$direct`$), thereby
increasing stability ($r_\delta = `r rich_on_stab$via_sync`$). On the other
hand, species richness had a positive direct effect on $\overline{CV_{sp}}$
($r_\delta = `r rich_on_cv_sp$direct`$),
thereby decreasing biomass stability ($r_\delta = `r rich_on_stab$via_cv_sp`$, Table
\@ref(tab:indir-sem)). The negative effect of $\overline{CV_{sp}}$ on temporal
stability was twice higher than the negative effect of synchrony (resp.
$r_\delta = `r cv_sp_on_stab`$ and $r_\delta = `r sync_on_stab`$, Fig.
\@ref(fig:sem-plot)D).

The total effect of mean trophic level and species richness on temporal
stability of biomass were of the same magnitude (resp.
$`r tlvl_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).
The mean trophic level had only a direct negative effect on
$\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$,
Fig. \@ref(fig:sem-plot)D) and thereby had a total positive
effect on temporal stability ($r_\delta = `r tlvl_on_stab$via_cv_sp["t_lvl"]`$,
Table \@ref(tab:indir-sem)). Connectance had no significant direct effect on either
$\overline{CV_{sp}}$ nor synchrony and thus no effect on temporal stability
(Fig. \@ref(fig:sem-plot)D, Table \@ref(tab:indir-sem)).

>>>>

---
title: Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages 
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault, Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
csl: ecology-letters.csl
---

@@

Biodiversity-ecosystem functioning (BEF) and food-web complexity-stability
relationships are central to ecology. However, they remain largely untested in
natural context. Here we study the links among environmental conditions,
richness, food-web structure, annual biomass and its temporal stability using a
standardized monitoring dataset of 99 stream fish communities spanning from 1995
to 2018. We first reveal that both richness and average trophic level are
positively related to annual biomass, with effects of similar strength. Second,
we show that food-web structure affects community stability, with a positive
effect of mean trophic level, and contrary to expectation, a destabilizing
effect of species richness. Finally, we find that environmental conditions
affect both biomass and its stability via effects on richness and network
structure. Strikingly, effects on stability are mediated by population stability
rather than synchrony, which contrasts with results from single-trophic
communities and might be characteristic of multi-trophic communities.

@@

Species richness has been found to increase primary productivity, community biomass
and its temporal stability in a diversity of ecosystems and taxa both
empirically and experimentally [@tilman_productivity_1996;
@cardinale_species_2002; @tilman_biodiversity_2006; @franssen_annual_2011;
@grace_integrative_2016; @pennekamp_biodiversity_2018; @houlahan_negative_2018;
@olivier_urbanization_2020]. However, the relationship between
diversity and ecosystem functioning remains mostly focused on single trophic
levels, generally the primary producers, ignoring the effects of food-web
structure. This gap is even more striking for the relationship between diversity
and temporal stability, in spite of long-standing interest for the links between
food-web complexity and stability in theoretical ecology [e.g.  @may_will_1972;
@dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011].

Overall, the evidence from theoretical models suggests that both species
richness and food-web structure should modulate ecosystem properties such as
community biomass and its stability. Complementarity in the use of
resources is important to understand the effects of species richness on
productivity in food-webs [@thebault_food-web_2003; @poisot_trophic_2013;
@schneider_animal_2016], but it also depends on food-webs connectance, i.e. its
level of generalism [@thebault_food-web_2003; @thebault_cascading_2007]. In
addition, the presence of species at higher trophic levels may increase total
primary production, through possibly through the selection of primary producers
which are larger and more complementary [@wang_biodiversity_2018].

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020],
asynchrony in population temporal variations also explains a significant part
of the stabilizing effects of species richness on community biomass in food-webs
[@thebault_trophic_2005]. Indeed, a low synchrony in population fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a population is compensated by the increase of another. However,
higher food-web connectance can counterbalance the positive effects of species
richness on temporal stability of community biomass, because it may increase
competition between consumers and then leads to a positive relationship between
species richness and population temporal variability
[@thebault_trophic_2005]. The addition of predators has
also cascading effects on both population variability and
synchrony at lower trophic levels [@teng_dynamics_2004;
@shanafelt_stability_2018]. Overall, these theoretical studies show that the
effects of species richness and food-web structure, such as connectance and
average trophic level, are potentially conflicting or synergistic in such a way
that they need to be addressed jointly.

The lack of studies that consider jointly effects of diversity and food-web
structure on ecosystem functions can be explained by the challenges associated
with manipulating both species richness and food-web complexity in an
experimental context [e.g. @sander_ecological_2017] and by the fact that
describing trophic interactions between species is highly labour intensive.
These challenges are even more acute for the study of the stability of ecosystem
properties, which additionally requires longer-term perspectives. In this
context, recent developments to predict species trophic interactions
[@beckerman_foraging_2006; @gravel_inferring_2013; @portalier_mechanics_2019],
combined with observational long-term surveys of ecological communities, can
offer the opportunity to investigate the links between diversity, food-web
structure and the stability of total community biomass in natural settings.

The joint study of the effects of diversity and food web structure on total
biomass and its stability is even more important in the context of environmental
changes. Recent studies on natural communities have shown that environment may
alter community biomass [@grace_integrative_2016; @woods_testing_2020] and its
temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus community biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
food-web structure, as larger ecosystems host food-webs with higher diversity
and trophic levels [@mchugh_dual_2010] whereas temperature can affect food-web
diversity and structure in contrasted ways depending on the ecosystem studied
[@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. In natural communities,
the relationships among species richness, community biomass and its temporal
stability might also arise from effects of the environment on both diversity and
ecosystem functions [@loreau_biodiversity_2001; @he_global_2013;
@grace_integrative_2016]. Thus, the effect of environment on ecosystem
functioning might arise from direct effects or indirectly through effects on
species richness or food-web structure.

@@

Because their trophic interactions are well known, stream fish communities
constitute a good model to study the effects of network structure on total
biomass and its temporal stability. A substantial body of literature
supports that trophic interactions can be inferred through body size and diet
informations [@jennings_weak_2001; @beckerman_foraging_2006;
@gravel_inferring_2013; @brose_predator_2019; @portalier_mechanics_2019].

@@

@@

We selected the stations that had been sampled ten years or more. As the concept
of community stability often refersred to the variability around an equilibrium
point [@donohue_navigating_2016], which is rarely observed or difficult to
characterize in natural communities. Then, we selected the stations that did not
show temporal trends in community biomass (detailed in appendix B, part 2).
However, the results were robust to the relaxation of this hypothesis (Fig. S7).
We selected the stations that had been sampled ten years or more. As the concept
of community stability referred to variability around an equilibrium, which is
rarely observed or difficult to characterize in natural communities, only those
stations that did not show temporal trends in community biomass were selected
(detailed in appendix B, part 2). However, the results were robust to the
relaxation of this hypothesis (Fig. S7). These selection steps resulted in 99
stations distributed over seven hydrographic basins (Fig. \@ref(fig:matmet)A).

@@

The body size of all fish individuals from a given species was not recorded when they were too
numerous (Fig. S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution (Fig. S1, appendix B, part 1.2). The body mass of each
individual was then estimated following the results of two meta-analyses on
fishes [@lleonart_removing_2000; @froese_cube_2006], using an allometric law
$B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams, $l$ the body
size in millimeters and $i$ the individual. Population biomass and community biomass at
each sampling event were then calculated by summing individual body masses. To
account for the variation in sampling effort among sites, population biomass and
community biomass were corrected by the surface sampled (range `r range_surface[1]` -
`r range_surface[2]` square metres), and expressed in grams per square meter
($gm^{-2}$). In the analysis, the community biomass of a site was defined as
the median of the community biomass across sampling years.

@@

Temporal biomass stability was defined as the inverse of biomass variability.
Variability was computed as Coefficient of Variation (CV) of the community biomass over
time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and $\mu_i$ being
respectively the standard deviation and the temporal average of the annual
community biomass of the station $i$). Then, temporal stability ($S_i$) was
equal to $CV_i^{-1} = \mu_i \sigma_i^{-1}$.

The variability of community biomass ($CV_{com, i}$) can be decomposed 
into a
synchrony ($\phi$) component and a weighted average population variability
($\overline{CV_{sp, i}}$) component according to @thibaut_understanding_2013:

@@

To assess the mechanisms driving stability of community biomass, we thus in addition measure
the synchrony and the weighted average population variability $\overline{CV_{sp,
i}}$, the average CV at the species-level in the station $i$, weighted by the
relative biomass of the species, $B_{k,i}$.

Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
community biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_{k,i}}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

@@

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all fishes and the resources (Fig.
SXX), then (2) determine network of a given community by
extracting the metaweb according to the fishes present at a given
site and sampling event (Fig. \@ref(fig:matmet)C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, larvae and small juveniles feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (appendix B, Fig. S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Fig. \@ref(fig:matmet)C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

@@

For each food-web, corresponding to a particular site and year, we computed the
connectance and mean trophic level. Connectance describes the level of
generalism in the network and is calculated as the actual number of interactions
divided by the number of interactions if all species interact ($C = L / N^2$, L
and N being resp. the number of links and of nodes).  The trophic level of the
nodes was computed as one added to the average trophic level of its food item.
The average trophic level $\overline{T}$ was computed as the average trophic
level of the trophic species weighted by their biomass (excluding resource nodes
for which we had no biomass information), so $\overline{T} = \sum_i^N{B_iT_i
\times (\sum_i^N{B_i})^{-1}}$, with $T_i$ and $B_i$ being respectively the
trophic level and the biomass of the node $i$. For each site, the median of
connectance and average trophic level over time was used in subsequent analysis
(Fig. \@ref(fig:matmet)C & D).

@@

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and temporal CV values of water temperature, flow, BOD, river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (appendix B, Fig. SXX). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Fig. \@ref(fig:matmet)B). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, average BOD and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

@@

We analyzed the relationships among environment, species richness, food-web
structure, community biomass and its temporal stability with structural equation
models as they enable the assessment of potential indirect and direct causal
relationships among multiple variables [@grace_structural_2008;
@grace_causal_2014; @lefcheck_piecewisesem_2016; @grace_integrative_2016].  We
defined a structural equation model explaining the median annual community
biomass as a function of the two synthetic environmental variables, species
richness and food-web structure [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and food-web
structure directly affect community biomass (Fig. \@ref(fig:sem-bm)A) and that
species richness affected food-web structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

We used the same model structure to assess the drivers of temporal stability
of community biomass (Fig.  \@ref(fig:sem-stab)A). Instead of the
annual community biomass, we modelled the synchrony and $\overline{CV_{sp}}$ which
in turn determined the temporal stability of community biomass. We linearized
the mathematical relationship linking temporal stability to $\overline{CV_{sp,
i}}$ and $\phi$ by taking the log of
temporal stability such as

\begin{equation} 
    \log S_i = - \log \overline{CV_{sp, i}} - \frac{1}{2} \log \phi_i
  (\#eq:stab)
\end{equation}

@@

The structural equation models explained a large part of the variation
of species richness, connectance, synchrony, $\overline{CV_{sp}}$ and total
biomass (resp. $R^2 = .43$, $.38$, $.46$, $.47$ and $.32$) but less so for mean
trophic level ($R^2 = .11$, Fig. \@ref(fig:sem-bm)A and \@ref(fig:sem-stab)A). We found a
negative relationship between species richness and mean trophic level ($r_\delta
= `r rich_on_tlvl`$, $r_\delta$ meaning standardized slope) and connectance
($r_\delta = `r rich_on_ct`$, Fig.  \@ref(fig:sem-bm)A and \@ref(fig:sem-stab)A), meaning that a
part of the effects of species richness on community biomass and its temporal
stability went through food-web structure.

@@

Species richness and mean trophic level had a significant positive direct effect
on community biomass ($r_\delta = `r rich_on_bm$direct`$ and 
$r_\delta = `r tlvl_on_bm`$ resp., Fig. \@ref(fig:sem-bm)A, Table
\@ref(tab:indir-sem)), while connectance had no direct effect.

The two PCA axes synthesising environmental variables had no significant direct
effects on community biomass, but indirect (Table \@ref(tab:indir-sem)). The first
axis, related to the stream size, was positively linked to community biomass
($r_\delta = `r env_on_bm$total["log_RC1"]`$, Table \@ref(tab:indir-sem)),
through both species richness and mean trophic level ($r_\delta = 
`r env_on_bm$via_richness["log_RC1"]`$ and $`r env_on_bm$via_tlvl["log_RC1"]`$
resp., Table \@ref(tab:indir-sem)). The second axis, related to the average
annual temperature was positively related
to community biomass ($r_\delta = `r env_on_bm$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)), mainly through species richness ($r_\delta = 
`r env_on_bm$via_richness["log_RC2"]`$, Table \@ref(tab:indir-sem)). Streams
with higher temperature and BOD contained more species, indirectly increasing
community biomass. The total effect of PCA1, PCA2 and species richness on
community biomass were of the same magnitude.

@@

By its effects on synchrony and population variability ($\overline{CV_{sp}}$), species richness had a
negative total effect on biomass stability ($r_\delta = `r rich_on_stab$total`$,
Table \@ref(tab:indir-sem)). This effect resulted from two opposite effects. On
the one hand, species richness had a direct negative effect on synchrony
($r_\delta = `r rich_on_sync$direct`$), thereby
increasing stability ($r_\delta = `r rich_on_stab$via_sync`$). On the other
hand, species richness had a positive direct effect on $\overline{CV_{sp}}$
($r_\delta = `r rich_on_cv_sp$direct`$),
thereby decreasing biomass stability ($r_\delta = `r rich_on_stab$via_cv_sp`$, Table
\@ref(tab:indir-sem)), which outweighted the decreasing effect of species
richness on synchrony. The
negative effect of $\overline{CV_{sp}}$ on temporal stability was twice higher
than the negative effect of synchrony (resp.
$r_\delta = `r cv_sp_on_stab`$ and $r_\delta = `r sync_on_stab`$, Fig.
\@ref(fig:sem-stab)A).

The total effect of mean trophic level and species richness on temporal
stability of biomass were of the same magnitude (resp.
$`r tlvl_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).
The mean trophic level had only a direct negative effect on
$\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$,
Fig. \@ref(fig:sem-stab)A) and thereby had a total positive
effect on temporal stability ($r_\delta = `r tlvl_on_stab$via_cv_sp["t_lvl"]`$,
Table \@ref(tab:indir-sem)). Food webs in which community biomass is located on
average at higher trophic levels thus tend to have higher population and
community stability. Connectance had no significant direct effect on either
$\overline{CV_{sp}}$ nor synchrony and thus no effect on temporal stability
(Fig. \@ref(fig:sem-stab)A, Table \@ref(tab:indir-sem)).
